% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[twocolumn]{article}
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage[english]{babel}

\input{preamble.tex}

\title{Student misconceptions in programming through the lens of variation 
theory}
\author{Celina Soori and Daniel Bosk (at the moment)}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed 

\begin{document}
\maketitle
\tableofcontents

\section{Introduction}

We study misconceptions in introductory programming from the perspective of 
variation theory.
We first survey the existing literature on misconceptions in introductory 
programming.
We then analyze these misconceptions using variation theory.

According to variation theory \parencite[Ch.~2]{NCOL}, each educational 
objective can be divided into different \emph{aspects}.
Consider the educational objective \enquote{the student should be able to use 
functions}.
One aspect of this particular educational objective is local and global scope 
of variables.
Another aspect is returning values from a function.
For a student to achieve the educational objective, she must be able to discern 
the different aspects.
Aspects that the student hasn't discerned are critical aspects.
One necessary condition for learning is that the student is introduced to a 
series of patterns of variation in the critical aspects.
Misconceptions occur when a student has failed to discern (at least) one 
critical aspect.

\paragraph{Purpose}

The purpose of this article is to give an overview of the studies that has been conducted on this subject and analyze these with help of variation theory to understand how one is supposed to teach CS1 to help students grasp important concepts in learning programming. Another purpose with this article is to discuss what further studies we need to conduct to understand deeper what students difficulties there are and how the education can be designed to avoid these. 

We want to do this as a base for future work using variation theory 
\parencite{NCOL}.

\section{Method}

I'm thinking we need a method-section to describe how this research has been conducted. Something about the litterature collecting and so on. 

\section{Concepts}

Added this to explain different words we use in the text and how we interpreted them and use them. Maybe we dont need subsections for every word, but I started this way.

\subsection{Variation Theory}

Here I'm thinking we should describe what variation theory is so that we can use it in the analysis later on.

\subsection{Misconception}

Do we need to describe what we mean with misconception? Do we maybe want to use another word? 

\subsection{Computational thinking}

Maybe we want to include something about computational thinking in this article? 

\subsection{Conceptual Change Theory}

An interesting method to unlearn misconceptions and instead learn the correct conception. Mentioned by Qian \& Lehman. 

\section{Important modules in CS1}

This section is divided to reflect the different concepts that are teached during CS1. Each section will describe what students often are meant to learn and understand in that module, which is then followed with a summary of what different studies have found is difficult for students in that module. 

\input{functions-variables.tex}
\input{classes.tex}
\input{repetitions.tex}
\input{types.tex}
\input{problem-solving.tex}
\input{conditionals.tex}

\subsection{Things that doesn't fit in anywhere else}

\input{tools.tex}


\section{Analysis}

The analyze will be performed with the help of variation theory. 

\section{Discussion}

Here we can discuss what we want to analyze and research further. We might also want to discuss how to design the education to avoid everything above.

\newpage

\begin{thebibliography}{9}

\bibitem{texbook}
Altadmri, A. Brown, N. C. C. (2015). 37 Million Compilations: Investigating Novice Programming Mistakes in Large-Scale Student Data. \emph{SIGCSE '15: Proceedings of the 46th ACM Technical Symposium on Computer Science Education}. February 2015. Pages 522-527.
\url{https://dl.acm.org/doi/abs/10.1145/2676723.2677258}

\bibitem{texbook}
Fleury, A. E. (1991). Parameter passing: The rules the students construct. \emph{SIGCSE’91:  In Proceedings of the 22nd SIGCSE Technical Symposium on Computer Science Education}. Pages 283-286.
\url{http://dx.doi.org/10.1145/107004.107066}

\bibitem{texbook}
Goldman, K. Gross, P. Heeren, C. Herman, G. Kaczmarczyk, L. Loui, M. C. Zilles, C. (2010). Identifying Important and Difficult Concepts in Introductory Computing Courses using a Delphi Process. \emph{SIGCSE '08: Proceedings of the 39th SIGCSE technical symposium on Computer science education} March 2008. Pages 256–260.
\url{https://dl.acm.org/doi/10.1145/1352135.1352226} 

\bibitem{texbook}
Holland, S. Griffiths, R. Woodman, M. (1997). Avoiding Object Misconceptions. \emph{ACM SIGCSE Bulletin} Volume 29, Issue 1, March 1997. Pages 131–134.
\url{https://dl.acm.org/doi/pdf/10.1145/268084.268132} 

\bibitem{texbook}
Kohn, T.  (2017). \emph{Variable Evaluation: an Exploration of Novice Programmers’ Understanding and Common Misconceptions}
\url{https://dl.acm.org/doi/abs/10.1145/3017680.3017724}

\bibitem{texbook}
Kurvinen, E. Hellgren, N. Kaila, E. Laakso, M. J. Salakoski, T. (2016). Programming misconceptions in an introductory level programming course exam. \emph{ITiCSE '16: Proceedings of the 2016 ACM Conference on Innovation and Technology in Computer Science Education}. July 2016. Pages 308–313. 
\url{https://doi.org/10.1145/2899415.2899447}

\bibitem{texbook}
Kumar Veerasamy, A. D’Souza, D. Laakso, M. J. (2016). Identifying Novice Student Programming Misconceptions and Errors From Summative Assessments. \emph{Journal of Educational Technology Systems.} Vol 45, Issue 1.  
\url{https://doi.org/10.1177/0047239515627263}

\bibitem{texbook}
Plass-Oude Bos, D.  (2015). Identifying and Addressing Common Programming Misconceptions with Variables (Part 1). \emph{University of Twente}.
\url{https://essay.utwente.nl/70455/1/Oude\%20Bos\%20Danny\%20-\%20S0021407\%20-\%20Afstudeerscriptie.pdf}

\bibitem{texbook}
Ragonis, N. Ben-Ari, M. (2005). A long-term investigation of the comprehension of OOP concepts by novices. \emph{ISSN: 0899-3408 (Print) 1744-5175 (Online) Journal homepage}. 15:3. 203-221. 
\url{https://doi.org/10.1080/08993400500224310}


\bibitem{texbook}
Sekiya Y. Yamaguchi, K. (2013) Tracing quiz set to identify novices' programming misconceptions. \emph{Koli Calling '13: Proceedings of the 13th Koli Calling International Conference on Computing Education Research} November 2013. Pages 87–95.
\url{https://doi.org/10.1145/2526968.2526978} 

\bibitem{texbook}
Sleeman, D. Putnam, R. T. Baxter, J. A. Kuspa, L. K. (1984). Pascal and High-School Students: A Study of Misconceptions. \emph{Technology Panel Study of Stanford and the Schools}. Occasional report \#009. 
\url{https://files.eric.ed.gov/fulltext/ED258552.pdf}

\bibitem{texbook}
Qian, Y. Lehman, J. (2017). Students’ Misconceptions and Other Difficulties in Introductory Programming: A Literature Review. \emph{ACM Trans. Comput. Educ.} 18, 1, Article 1 (October 2017).
\url{https://doi.org/10.1145/3077618} 



\end{thebibliography}
\printbibliography

\end{document}
