\subsection{Problem solving}

\subsubsection{Role in the syllabus}

\subsubsection{Difficulties that can occur}


Here I want to have some articles about how math-problems will make it harder for students to solve the problem. Quote from Veerasamy et al \emph{This study analysis also explored that novices of programming struggled in writing code for math-related Questions 6 and 7 (refer Table C1). Nearly 66\% of students did not do well in the mathematical problem-based questions though explained and allowed to surf the Internet to seek for more details during the exam hours. A neo-Piagetian theory of cognitive development stated that students who are at the concrete operational stage struggle to write large programs with partial specifications, although they can write small programs from well-defined specifications (Teague et al., 2012).}

Also I would want to include difficulties students have when debugging the code and trying to find errors. Students often have a problem with tracing the code, something that is discussed in page 20 by Sleeman et al., 1984. 
On the same subject as above: In what order a program will be executed in, Programming misconceptions in an introductory level programming course exam by Einari Kurvinen, Niko Hellgren, Erkki Kaila, Mikko-Jussi Laakso, Tapio Salakoski

Would also maybe like to mention how a lab instruction should be to help students get the right knowledge from the lab. Is discussed somewhere in Yizhou Qian and James Lehmans article I think.

