\subsection{Repetitions}

In CS1 students usually learn about repetitions, which includes for- and while
-loops and in some cases recursion.

Loop constructions can be hard to trace and understand for novice students, 
for instance when a loop starts, ends and what is repeated and not repeated 
in the loop \parencite{Sekiya2013,KumarVeerasamy2016,Kaczmarczyk2010}. This 
was something \textcite{Sleeman1984} also realised in their article when 
studying high school students writing and debugging loop-structures. A common 
misconception that the students had was that if the loop contained a print-
statement, the students thought that the only thing  repeated inside the loop 
was the string they saw in the terminal. The difficulties students have in 
tracing the code linearly when entering a loop is according to \textcite{
KumarVeerasamy2016} because of the lack of understanding the students have of 
the looping technique and the amount of cognitive skills the tracing takes.

XXX Add analysis on how we can help students trace loops and understanding 
how the loop-structure works.

Another difficult part of the loop technique is to understand how an if-
statement inside a loop is executed. \textcite{Sekiya2013} found in their 
studies that the combination of the two control structures created 
misconceptions. For instance the students thought that the variables in the 
conditional part of the loop-construction was control variables or the output 
from the loop. The students in the studies often got confused and started to 
misplace the different variables that are defined when writing an if-
statement in a for-loop.

XXX Add analysis on how we can teach the combination of loops and 
conditionals in a way which will avoid misconceptions about the different 
variables used. 


\textbf{Fun things to add:}

- "Singular/plural: A common idiom for iteration is to use a plural noun for 
a collection and singular for each element in it, such as ‘for name in names: 
<loop body>’. Some learners were confused about whether to access elements 
inside the for-loop using singular or plural. Singular name is correct most 
of the time, but sometimes the entire collection needs to be accessed, such 
as names.remove()." \parencite{GuoMarkelZhang2020}

- "Iteration that reads like English: Variables like i are often used to 
iterate from 0 to some upper bound. Some learners wrote code like ‘while i <= 
100’ or ‘for i in 100’, which read in English like i will iterate up to 100. 
The former fails because i is never initialized, and the latter fails because 
for-loops need a collection to iterate over." \parencite{GuoMarkelZhang2020}